---
title: "Homework 6"
author: "Ekaterina Hofrenning"
date: '2023-11-27'
output: github_document
---

```{r, message = F, warning = F}
library(tidyverse)
library(rnoaa)
library(janitor)
library(MASS)
library(modelr)
set.seed(254)
```

# Problem 1

```
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```

# Problem 2

```
The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:

- r^2
- log(b1*b2)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(b1*b2).
```

First, I read in the data and clean up some formatting issues.

```{r, message = F, warning = F}
# read data
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  dplyr::select(name, id, everything())
```

Next, I run the 5000 bootstrapped models and extract the elements of interest.

```{r, message = F, warning = F}
# run the bootstraps
bootstrap_df =
  weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    rsquared = map(models, broom::glance)) %>%
  dplyr::select(results, rsquared) %>% unnest(rsquared) %>%
  dplyr::select(results, r.squared) %>% unnest(results) %>%
  dplyr::select(term, estimate, r.squared) 
```

Below, I plot the R^2 terms and calculate a CI for these values.

```{r, message = F, warning=F}
# R^2 plot
bootstrap_df %>%
  filter(term == "(Intercept)") %>%
  ggplot(aes(x = r.squared)) +
  geom_density()

# CI for R^2
r2_CI <-
  bootstrap_df %>%
  filter(term == "(Intercept)") %>%
  dplyr::select(r.squared) %>%
  summarize(ci_lower = quantile(r.squared, 0.025), 
            ci_upper = quantile(r.squared, 0.975))

r2_CI
```

The distribution of the 5000 R-squared values has a slight left skew and is centered near .92. The 95% confidence interval of the R-squared values is between: `r r2_CI$ci_lower` and `r r2_CI$ci_upper`.


Below, I plot the **log(beta1*beta2)** estimates and calculate a confidence interval similar to above.

```{r, warning = F, message = F}
# plot log(b1*b2)
bootstrap_df %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
  rename(beta_1 = tmin, 
         beta_2 = prcp) %>% 
  mutate(log_b1_b2 = log(beta_1 * beta_2)) %>% 
  ggplot(aes(x = log_b1_b2)) + 
  geom_density()

# CI for log(b1*b2)
log_b1b2_CI <-
bootstrap_df %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
  rename(beta_1 = tmin, 
         beta_2 = prcp) %>% 
  mutate(log_b1_b2 = log(beta_1 * beta_2)) %>%
  summarize(NA_num = sum(is.na(log_b1_b2)),
            ci_lower = quantile(log_b1_b2, 0.025, na.rm = T), 
            ci_upper = quantile(log_b1_b2, 0.975, na.rm = T))

log_b1b2_CI
```

The distribution of the 5000 **log(b1*b2)** values has a large left skew. It's peak is around -5. When we take the log of negative values, NAs are produced. In this case, `r log_b1b2_CI$NA_num` NAs were produced. Of those that are non-NA, the confidence interval for **log(b1*b2)** is between: `r log_b1b2_CI$ci_lower` and `r log_b1b2_CI$ci_upper`.


# Problem 3

```
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```

```{r, message = F}
bw <- read_csv("data/birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(frace, 
                       "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4", "other" = "8"),
    malform = as.logical(malform),
    mrace = as.factor(mrace),
    mrace = fct_recode(mrace, "white" = "1", "black" = "2", 
                       "asian" = "3", "puerto rican" = "4")) %>%
  drop_na()
```

```
Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.
```

In order to choose model features, I will use a backward stepwise selection algorithm from the MASS package. This uses the Aikake's Information Criteria (AIC) measure to kick out predictors one by one in order to find the most parsimonious model.

```{r}
# My model (1)
lm1 <- lm(bwt ~., data = bw)
step <- stepAIC(lm1, direction = "backward", trace = FALSE)
step %>% broom::tidy() %>% knitr::kable()
```


Here is a residual plot for my model (residuals vs. fitted values):

```{r}
bw %>%  
  add_predictions(step) %>% 
  add_residuals(step) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  labs(title = "Residuals vs. Fitted Values ", 
       x = "Fitted",
       y = "Residuals")
```

```
Compare your model to two others:

- One using length at birth and gestational age as predictors (main effects only)
- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```

Initialize Comparison Models:

```{r}
# Comparison model (2)
lm2 <- lm(bwt ~ blength + gaweeks, data = bw)
lm2 %>% broom::tidy() %>% knitr::kable()

# Comparison model (3)
lm3 <- lm(bwt ~ bhead*blength*babysex, data = bw)
lm3 %>% broom::tidy() %>% knitr::kable()
```

CV Prediction Error:

Numerically showing the RMSE for each model:

```{r, warning = F, message = F}
cv_df <-
  crossv_mc(bw, 100) %>% 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble))

cv_comp <-
cv_df %>% 
  mutate(
    mod_mine  = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + fincome + 
    gaweeks + mheight + mrace + parity + ppwt + smoken, data = .x)),
    mod2  = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
    mod3  = map(train, ~lm(bwt ~ bhead*blength*babysex, data = .x))) %>% 
  mutate(
    rmse_mine = map2_dbl(mod_mine, test, ~rmse(model = .x, data = .y)),
    rmse_2  = map2_dbl(mod2, test, ~rmse(model = .x, data = .y)),
    rmse_3 = map2_dbl(mod3, test, ~rmse(model = .x, data = .y))) %>% 
  summarize(
    avg_rmse_mod_mine = mean(rmse_mine),
    avg_rmse_mod2 = mean(rmse_2),
    avg_rmse_mod3 = mean(rmse_3))

cv_comp
```

Plotting RMSE for each model:
```{r, warning = F}
cv_df %>% 
  mutate(mod_mine  = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + 
                                      fincome + gaweeks + mheight + mrace + parity + 
                                      ppwt + smoken, data = .x)),
         mod_comp1  = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
         mod_comp2  = map(train, ~lm(bwt ~ bhead*blength*babysex, data = .x))) %>% 
  mutate(rmse_mine = map2_dbl(mod_mine, test, ~rmse(model = .x, data = .y)),
         rmse_comp1  = map2_dbl(mod_comp1, test, ~rmse(model = .x, data = .y)),
         rmse_comp2 = map2_dbl(mod_comp2, test, ~rmse(model = .x, data = .y))) %>%
  dplyr::select(rmse_mine, rmse_comp1, rmse_comp2) %>%  
  pivot_longer(everything(),
               names_to = "model", 
               values_to = "rmse",
               names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin(aes(fill = model))
```

The average RMSE is lowest for my model (`r cv_comp %>% pull(avg_rmse_mod_mine)`)compared to the 2 models defined above, compared to `r cv_comp %>% pull(avg_rmse_mod2)` for the first comparison model and `r cv_comp %>% pull(avg_rmse_mod3)` for the second comparison model.

